---
title: "Assignment#3"
author: "Tyler Cobian"
date: "4/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load in the packages needed
```{r, include=FALSE}
library(tidyverse)
library(sensitivity)
library(gridExtra)
library(deSolve)
library(pse)

```
Adjust your almond model to  output ONLY the mean almond yield anomoly IF the users sets parameter (e.g mean_only = TRUE))

```{r}
# import the climate data
clim <- read.table("clim.txt",  sep=" ", header=T)

almond_yield <- function(mean_only,
                         clim_data = clim,
                       a=-0.015,
                       b=-0.0046,
                       c=-0.07,
                       d=0.0043,
                       e=0.28) {
  
  # Add in some error checking into the function
  
  # Make sure the climate data input is a dataframe
  if(class(clim_data) != "data.frame") return("Climate data input must be a data frame")
  
   # Make sure the climate data input contains the columns year, month, precip, tmin_c and tmax_c
  if(!all(has_name(clim_data, c("month", "year",  "month", "precip", "tmin_c", "tmax_c")))) return("Climate data input must contain the following columns: year, month, precip, tmin_c and tmax_c")
  
  # Make sure that the input for precipitation is larger than 0
  clim_data$precip = ifelse(clim_data$precip < 0, return("Input for precipitation must be a value larger than 0"), clim_data$precip)
  
  # Make sure that the maximum tempertaure will be larger than the minimum temperature
  clim_data$tmin_c = ifelse(clim_data$tmin_c > clim_data$tmax_c, return("Input for maximum temperature must be larger than input for minimum temperature"), clim_data$tmin_c)
  
  # Average monthly maximum daily temperature, and monthly precipitation from a data frame called clim  with columns year, month, precip and tmax_c

clim_month <-  clim_data %>%
group_by(month, year) %>%
dplyr::summarize(meantmin = mean(tmin_c),
          meantmax = mean(tmax_c),
          precip=sum(precip))
  
   # Filter Jan and Feb data
  jan <- clim_month %>% 
    filter(month==1)
  
  feb <- clim_month %>% 
    filter(month==2)
  
  # Change column names
  colnames(jan) <- c("month", "year", "Tn", "Tm", "P")
  colnames(feb) <- c("month", "year", "Tn", "Tm", "P")
  
  # Data structure for yield annomalies
  yield_df <- data.frame(year = jan$year, YA = NA)
  
  # Loop through each year
  for (i in 1:length(yield_df$year)) {
   yield_df$YA[i] = a*feb$Tn[i] + b*(feb$Tn[i]^2) +c*jan$P[i] + d*(jan$P[i]^2) + e
}
  
  # Calculate max and min yields
  max_yield <- yield_df %>% 
    arrange(-abs(YA)) %>% 
    head(1)
  
  min_yield <- yield_df %>% 
    arrange(abs(YA)) %>% 
    head(1)
  
  mean_YA<- mean(yield_df$YA)
  
  # Change column names of max and min yields
  colnames(max_yield) <- c("Year", "Maximum Yield Anomaly")
  colnames(min_yield) <- c("Year", "Minimum Yield Anomaly")
  
  # Create list with three elements
  yield_list <- list(yield_df, max_yield, min_yield)
  
 if(mean_only)
    return(mean(yield_df$YA))
 else
    return(yield_list)

}
  
almond_yield(mean_only = T)

```
LHS Sensativity of all the parameters

```{r}
# Set up the parameters to test
factors = c("a", "b", "c", "d", "e")

# Set the number of parmeter sets to run
nsets = 100

# make the distribution of th parameters
q = c("qnorm","qnorm","qnorm","qnorm","qnorm")

q.arg = list(list(mean = 0.015, sd = 0.015*0.20), list(mean = 0.0046, sd = 0.0046*0.20), list(mean = 0.07, sd = 0.07*0.2), list(mean = 0.0043, sd = 0.0043*0.20), list(mean = 0.28, sd = 0.28*0.20))

# Generate samples from LHS
sen_yield_LHS<- LHS(NULL, factors, nsets, q, q.arg)
sen_yield_LHS_pars<- get.data(sen_yield_LHS)
head(sen_yield_LHS_pars)

# Create a data structure to store the resulst of running these parameter values
LHS_sen_results<- matrix(nrow = nsets, ncol = 1)

# Use a for loop to run these new parameter values through the alomd_yield function
for(i in 1:nsets){
  LHS_sen_results[i,] = unlist(almond_yield(clim_data = clim, mean_only = T,
                                            a = sen_yield_LHS_pars[i,1],
                                            b = sen_yield_LHS_pars[i,2],
                                            c = sen_yield_LHS_pars[i,3],
                                            d = sen_yield_LHS_pars[i,4],
                                            e = sen_yield_LHS_pars[i,5]))
}

# Clean up the results data frame
colnames(LHS_sen_results)<- "avg_yield"

LHS_sen_results = as.data.frame(LHS_sen_results)

# look at the senativity of each parameter
sen_yield_LHS<- pse::tell(sen_yield_LHS, t(LHS_sen_results))
sen_yield_LHS

```
Rank the parameters in term of their sensitivity
Graph uncertainty in mean yield anomaly across all parameter uncertainty (boxplot and cummulative distribution of the output).

### Sensativity of Parameters in order of most sensative
#### 1. D
#### 2. C
#### 3. B
#### 4. A
#### 5. E

```{r}
# Make a boxplot of the LHS results
LHS_plot<- ggplot(data = LHS_sen_results, aes(y = avg_yield, colour = avg_yield))+
  geom_boxplot()+
  labs(y = "Avgerage Yearly Yield", title = "Sensativity of Yearly Yields on Model Parameters")+
  theme_minimal()+
  theme(axis.title.y = element_blank())
LHS_plot
```
Repeat using Sobel method

```{r}
# Sobel method for sensativity

# Set the number of patameter sets
np = 100

# Create distributions for the different paramenters

a_dist_1<- rnorm(mean = 0.015, sd = 0.015*0.20, n = np)
b_dist_1<- rnorm(mean = 0.0046, sd = 0.0046*0.20, n = np)
c_dist_1<- rnorm(mean = 0.07, sd = 0.07*0.20, n = np)
d_dist_1<- rnorm(mean = 0.0043, sd = 0.0043*0.20, n = np)
e_dist_1<- rnorm(mean = 0.28, sd = 0.28*0.20, n = np) 


a_dist_2<- rnorm(mean = 0.015, sd = 0.015*0.20, n = np)
b_dist_2<- rnorm(mean = 0.0046, sd = 0.0046*0.20, n = np)
c_dist_2<- rnorm(mean = 0.07, sd = 0.07*0.20, n = np)
d_dist_2<- rnorm(mean = 0.0043, sd = 0.0043*0.20, n = np)
e_dist_2<- rnorm(mean = 0.28, sd = 0.28*0.20, n = np) 
# make a data set with these new parameters 

sens_1<- cbind.data.frame(a = a_dist_1, b = b_dist_1, c = c_dist_1, d = d_dist_1, e = e_dist_1)

sens_2<- cbind.data.frame(a = a_dist_2, b = b_dist_2, c = c_dist_2, d = d_dist_2, e = e_dist_2)

sen_yield<- sobol2007(model = NULL, sens_1, sens_2, nboot = 100)

# Run the model through the new parameters
run_1<- mapply(almond_yield,
               a = sen_yield$X$a,
               b = sen_yield$X$b,
               c = sen_yield$X$c,
               d = sen_yield$X$d,
               e = sen_yield$X$e,
               mean_only = T)

# make a data frame of run_1
run_1_data<- as.data.frame(run_1)

# give results to sensativity analysis
sen_yield<- sensitivity::tell(sen_yield, run_1)

# First order indicies (main effect without any co-variance)
sen_yield$S
# Total sensativity index
sen_yield$T

# Make a data frame of the total senastivites
sobol_sensitivity<- data.frame("parameter" = c("a", "b", "c", "d", "e"),
                               "original" = sen_yield$T$original,
                               "bias" = sen_yield$T$bias,
                               "std_err" = sen_yield$T$`std. error`,
                               "min_ci" = sen_yield$T$`min. c.i.`,
                               "max_ci" = sen_yield$T$`max. c.i.`)

# Vizualize the results
print(sen_yield)
plot(sen_yield)

```
Repeat using twice as many parameter sets as you did in your first sensitivity analysis - and look at how this changes the sensitivity results

```{r}
# repeat the LHS sensativity but with 500 runs
# Set up the parameters to test
factors = c("a", "b", "c", "d", "e")

# Set the number of parmeter sets to run
nsets = 500

# make the distribution of th parameters
q = c("qnorm","qnorm","qnorm","qnorm","qnorm")

q.arg = list(list(mean = 0.015, sd = 0.20), list(mean = 0.0046, sd = 0.20), list(mean = 0.07, sd = 0.2), list(mean = 0.0043, sd = 0.20), list(mean = 0.28, sd = 0.20))

# Generate samples from LHS
sen_yield_LHS<- LHS(NULL, factors, nsets, q, q.arg)
sen_yield_LHS_pars<- get.data(sen_yield_LHS)
head(sen_yiels_LHS_pars)

# Create a data structure to store the resulst of running these parameter values
LHS_sen_results<- matrix(nrow = nsets, ncol = 1)

# Use a for loop to run these new parameter values through the alomd_yield function
for(i in 1:nsets){
  LHS_sen_results[i,] = unlist(almond_yield(clim_data = clim, mean_only = T,
                                            a = sen_yield_LHS_pars[i,1],
                                            b = sen_yield_LHS_pars[i,2],
                                            c = sen_yield_LHS_pars[i,3],
                                            d = sen_yield_LHS_pars[i,4],
                                            e = sen_yield_LHS_pars[i,5]))
}

# Clean up the results data frame
colnames(LHS_sen_results)<- "avg_yield"

LHS_sen_results = as.data.frame(LHS_sen_results)

sen_yield_LHS<- pse::tell(sen_yield_LHS, t(LHS_sen_results))
sen_yield_LHS

```
Submit R markdown and short write up describing what you learned from the sensitivity analysis

The results of the sensativity analysis concluded that the "D" parameter of the almond yield function was most sensative followed by "C", "B", "A", "E". 








































