---
title: "Assignment#3"
author: "Tyler Cobian"
date: "4/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load in the packages needed
```{r, include=FALSE}
library(tidyverse)
library(sensitivity)
library(gridExtra)
library(deSolve)

```
Adjust your almond model to  output ONLY the mean almond yield anomoly IF the users sets parameter (e.g mean_only = TRUE))

```{r}
# import the climate data
clim <- read.table("clim.txt",  sep=" ", header=T)

almond_yield <- function(mean_only, 
                      clim_data = clim,
                       a=-0.015,
                       b=-0.0046,
                       c=-0.07,
                       d=0.0043,
                       e=0.28) {
  
  # Add in some error checking into the function
  
  # Make sure the climate data input is a dataframe
  if(class(clim_data) != "data.frame") return("Climate data input must be a data frame")
  
   # Make sure the climate data input contains the columns year, month, precip, tmin_c and tmax_c
  if(!all(has_name(clim_data, c("month", "year",  "month", "precip", "tmin_c", "tmax_c")))) return("Climate data input must contain the following columns: year, month, precip, tmin_c and tmax_c")
  
  # Make sure that the input for precipitation is larger than 0
  clim_data$precip = ifelse(clim_data$precip < 0, return("Input for precipitation must be a value larger than 0"), clim_data$precip)
  
  # Make sure that the maximum tempertaure will be larger than the minimum temperature
  clim_data$tmin_c = ifelse(clim_data$tmin_c > clim_data$tmax_c, return("Input for maximum temperature must be larger than input for minimum temperature"), clim_data$tmin_c)
  
  # Average monthly maximum daily temperature, and monthly precipitation from a data frame called clim  with columns year, month, precip and tmax_c

clim_month <-  clim_data %>%
group_by(month, year) %>%
summarize(meantmin = mean(tmin_c),
          meantmax = mean(tmax_c),
          precip=sum(precip))
  
   # Filter Jan and Feb data
  jan <- clim_month %>% 
    filter(month==1)
  
  feb <- clim_month %>% 
    filter(month==2)
  
  # Change column names
  colnames(jan) <- c("month", "year", "Tn", "Tm", "P")
  colnames(feb) <- c("month", "year", "Tn", "Tm", "P")
  
  # Data structure for yield annomalies
  yield_df <- data.frame(year = jan$year, YA = NA)
  
  # Loop through each year
  for (i in 1:length(yield_df$year)) {
   yield_df$YA[i] = a*feb$Tn[i] + b*(feb$Tn[i]^2) +c*jan$P[i] + d*(jan$P[i]^2) + e
}
  
  # Calculate max and min yields
  max_yield <- yield_df %>% 
    arrange(-abs(YA)) %>% 
    head(1)
  
  min_yield <- yield_df %>% 
    arrange(abs(YA)) %>% 
    head(1)
  
  mean_YA<- mean(yield_df$YA)
  
  # Change column names of max and min yields
  colnames(max_yield) <- c("Year", "Maximum Yield Anomaly")
  colnames(min_yield) <- c("Year", "Minimum Yield Anomaly")
  
  # Create list with three elements
  yield_list <- list(yield_df, max_yield, min_yield)
  
 if(mean_only)
    return(mean(yield_df$YA))
 else
    return(yield_list)

}
  
almond_yield(mean_only = T)

```
Sobol Sensativity of all the parameters

```{r}
# Sobel method for sensativity

# Set the number of patameter sets
np = 100

# Create distributions for the different paramenters

a_dist_1<- rnorm(mean = 0.015, sd = 0.20, n = np)
b_dist_1<- rnorm(mean = 0.0046, sd = 0.20, n = np)
c_dist_1<- rnorm(mean = 0.07, sd = 0.20, n = np)
d_dist_1<- rnorm(mean = 0.0043, sd = 0.20, n = np)
e_dist_1<- rnorm(mean = 0.28, sd = 0.20, n = np) 


a_dist_2<- rnorm(mean = 0.015, sd = 0.20, n = np)
b_dist_2<- rnorm(mean = 0.0046, sd = 0.20, n = np)
c_dist_2<- rnorm(mean = 0.07, sd = 0.20, n = np)
d_dist_2<- rnorm(mean = 0.0043, sd = 0.20, n = np)
e_dist_2<- rnorm(mean = 0.28, sd = 0.20, n = np) 
# make a data set with these new parameters 

sens_1<- cbind.data.frame(a = a_dist_1, b = b_dist_1, c = c_dist_1, d = d_dist_1, e = e_dist_1)

sens_2<- cbind.data.frame(a = a_dist_2, b = b_dist_2, c = c_dist_2, d = d_dist_2, e = e_dist_2)

sen_yield<- sobol2007(model = NULL, sens_1, sens_2, nboot = 100)

# Run the model through the new parameters
run_1<- mapply(almond_yield,
               a = sen_yield$X$a,
               b = sen_yield$X$b,
               c = sen_yield$X$c,
               d = sen_yield$X$d,
               e = sen_yield$X$e,
               mean_only = T)

# give results to sensativity analysis
sen_yield<- tell(sen_yield, run_1)

# First order indicies (main effect without any co-variance)
sen_yield$S
# Total sensativity index
sen_yield$T

# Vizualize the results
print(sen_yield)
plot(sen_yield)


```
Rank the parameters in term of their sensitivity
Graph uncertainty in mean yield anomaly across all parameter uncertainty (boxplot and cummulative distribution of the output).

```{r}

```
Repeat using the LHS and Sobel methods

```{r}

```
Repeat using twice as many parameter sets as you did in your first sensitivity analysis - and look at how this changes the sensitivity results

```{r}

```
Submit R markdown and short write up describing what you learned from the sensitivity analysis

```{r}

```







































